---
title: "Metformin LCA"
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    echo: false
    warning: false
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(kableExtra, dplyr, tidyr, lme4, lmerTest, ggplot2, cowplot,
               mixOmics, ggalluvial)
devtools::load_all()
```

```{r}
#| label: data-prep
#| include: false

data <- lca.diablo::met_analytic

# replicate previous LCA using full dataset, not just UCD subset with omics
data_lca_complete <- data |> 
  dplyr::mutate(
    visit = dplyr::case_when(
      redcap_event_name == "Baseline Visit (Visit 1)" ~ 1,
      redcap_event_name == "Week 8 Visit (Visit 2)" ~ 2,
      redcap_event_name == "Week 16 (Visit 3/Early Term)" ~ 3
    ),
    adams_total = dplyr::if_else(
      dplyr::if_all(
        .cols = c(adams_manic, adams_depress, adams_avoid, 
                  adams_anxious, adams_obsess), 
        ~ is.na(.x)
      ),
      NA,
      rowSums(
        dplyr::across(
          .cols = c(adams_manic, adams_depress, adams_avoid, 
                    adams_anxious, adams_obsess)
        ), 
        na.rm = TRUE
      )
    ),
    pql_total = dplyr::if_else(
      dplyr::if_all(
        .cols = c(pql_pf_tot, pql_ef_tot, pql_socf_tot, pql_schf_tot), 
        ~ is.na(.x)
      ),
      NA,
      rowSums(
        dplyr::across(
          .cols = c(pql_pf_tot, pql_ef_tot, pql_socf_tot, pql_schf_tot)
        ), 
        na.rm = TRUE
      )
    )
  ) |> 
  dplyr::select(
    fxs_sts_id, Treatment, age_group, visit, 
    # updated list using results from clinical trial
    abc_comp_rev, 
    adams_avoid, adams_total, 
    pql_schf_tot, pql_total,
    cshq_ts
  ) |> 
  # update names to match previous data/code
  dplyr::rename(
    id = fxs_sts_id,
    ADAMS_Total = adams_total,
    PedsQL = pql_total
  )
```


# LCA 

Fit LCA using the following clinical variables as outcomes: 
ABC Composite Score, ADAMS total score, and Peds QL total score.  

Fit LCA using k = 1, 2, and 3 and determine the optimal number of clusters 
based on AIC. Each k was iterated 10 times and the best run, 
maximum pseudo-likelihood, was selected to reduce the effect of the initial 
cluster assignment. 

::: {.callout-important}  

Initially tried to fit LCA using 1, 2, 3, and 4 clusters but Age 6-11 
continuously crashed when k = 4.  

:::

## LCA - All

```{r}
data_lca <- data_lca_complete
```

{{< include _sec-lca.qmd >}}

## LCA - Age 6-11

```{r}
data_lca <- data_lca_complete |> 
  dplyr::filter(age_group == "Age 6-11")
```

{{< include _sec-lca.qmd >}}

## LCA - Age 12+

```{r}
data_lca <- data_lca_complete |> 
  dplyr::filter(age_group == "Age 12+")
```

{{< include _sec-lca.qmd >}}
