---
title: "Metformin DIABLO"
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    echo: false
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(kableExtra, dplyr, tidyr, lme4, lmerTest, ggplot2, cowplot,
               mixOmics)
devtools::load_all()
```

```{r}
#| label: import-results
#| include: false

## load data ##
X_diff <- readRDS(
  file = "../extdata/archive/DIABLO_X_Difference_list.rds"
)
# update X_diff names
names(X_diff) <- c("Metabolites", "Lipids", "Peptides", "Proteomics")

# remove subj 10
X_diff$Metabolites <- X_diff$Metabolites[-10, ]
X_diff$Lipids <- X_diff$Lipids[-10, ]
X_diff$Peptides <- X_diff$Peptides[-10, ]
X_diff$Proteomics <- X_diff$Proteomics[-10, ]

Y <- readRDS(
  file = paste0(
    "../extdata/archive/k2.Assignments.20230714.rds"
  )
)

design.mat <- readRDS(
  file = "../extdata/archive/Metformin.Difference.Diablo.Design.Matrix.rds"
)

rownames(design.mat) <- colnames(design.mat) <- c(
  "Metabolites", "Lipids", "Peptides", "Proteomics"
)

## load lca results ##
metformin_lca_2 <- readRDS(
  paste0(
    "../extdata/archive/Metformin.k2.LCA.NoOutlier.rds"
  )
)

## load DIABLO results ##
k2.diablo.tune.final <- readRDS(
  paste0(
    "../extdata/archive/k2.DIABLO.Tuned.20230713.rds"
  )
)

## omics ##
omics_list <- readRDS(
  paste0(
    "../extdata/archive/DIABLO_X_list.rds"
  )
)

## visit designs ##
patient_design <- readRDS(
  paste0(
    "../extdata/archive/DiffExp_Design_20230105.rds"
  )
)
```

# Previous Analysis 

## Clustering

```{r avg-response-by-cluster}
lca_2_means <- data.frame(
  "Cluster" = as.factor(c(rep(c(1, 2), each = 3))),
  "Visit" = rep(c(1, 2, 3), 2),
  "ABC.Composite" = c(
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "abc_comp_rev"] * 1,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "abc_comp_rev"] * 2,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "abc_comp_rev"] * 3,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "abc_comp_rev"] * 1,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "abc_comp_rev"] * 2,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "abc_comp_rev"] * 3
  ),
  "ADAMS.Total" = c(
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "ADAMS_Total"] * 1,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "ADAMS_Total"] * 2,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "ADAMS_Total"] * 3,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "ADAMS_Total"] * 1,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "ADAMS_Total"] * 2,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "ADAMS_Total"] * 3
  ),
  "PedsQL" = c(
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "PedsQL"] * 1,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "PedsQL"] * 2,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "PedsQL"] * 3,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "PedsQL"] * 1,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "PedsQL"] * 2,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "PedsQL"] * 3
  )
) |>
  pivot_longer(cols = c(ABC.Composite, ADAMS.Total, PedsQL),
               names_to = "Measure", values_to = "Score") |>
  mutate(
    y_min = 0,
    y_max = case_when(
      Measure == "ABC.Composite" ~ 80,
      Measure == "ADAMS.Total" ~ 40,
      Measure == "PedsQL" ~ 300
    )
  )

# average Std Error
lca_2_ste <- data.frame(
  Cluster = c(1, 2),
  ABC.Composite = c(
    mean(
      c(metformin_lca_2$Models$`1`$Fixedest[[1]]["visit1_", "Std. Error"],
        metformin_lca_2$Models$`1`$Fixedest[[2]]["visit1_", "Std. Error"])
    ),
    mean(
      c(metformin_lca_2$Models$`2`$Fixedest[[1]]["visit1_", "Std. Error"],
        metformin_lca_2$Models$`2`$Fixedest[[2]]["visit1_", "Std. Error"])
    )
  ),
  ADAMS.Total = c(
    mean(
      c(metformin_lca_2$Models$`1`$Fixedest[[1]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`1`$Fixedest[[3]]["visit1_", "Std. Error"])
    ),
    mean(
      c(metformin_lca_2$Models$`2`$Fixedest[[1]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`2`$Fixedest[[3]]["visit1_", "Std. Error"])
    )
  ),
  PedsQL = c(
    mean(
      c(metformin_lca_2$Models$`1`$Fixedest[[2]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`1`$Fixedest[[3]]["visit2_", "Std. Error"])
    ),
    mean(
      c(metformin_lca_2$Models$`2`$Fixedest[[2]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`2`$Fixedest[[3]]["visit2_", "Std. Error"])
    )
  )
)

# add +/- 1.96*StdError
test <- lca_2_means |>
  mutate(
    seg.min = case_when(
      Measure == "ABC.Composite" & Cluster == 1 ~ Score - (1.96 * lca_2_ste["1", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 1 ~ Score - (1.96 * lca_2_ste["1", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 1 ~ Score - (1.96 * lca_2_ste["1", "PedsQL"]),
      Measure == "ABC.Composite" & Cluster == 2 ~ Score - (1.96 * lca_2_ste["2", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 2 ~ Score - (1.96 * lca_2_ste["2", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 2 ~ Score - (1.96 * lca_2_ste["2", "PedsQL"])
    ),
    seg.max = case_when(
      Measure == "ABC.Composite" & Cluster == 1 ~ Score + (1.96 * lca_2_ste["1", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 1 ~ Score + (1.96 * lca_2_ste["1", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 1 ~ Score + (1.96 * lca_2_ste["1", "PedsQL"]),
      Measure == "ABC.Composite" & Cluster == 2 ~ Score + (1.96 * lca_2_ste["2", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 2 ~ Score + (1.96 * lca_2_ste["2", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 2 ~ Score + (1.96 * lca_2_ste["2", "PedsQL"])
    )
  )

# rename variables for figures
test <- test |>
  mutate(
    Measure = case_when(
      Measure == "ABC.Composite" ~ "ABC Composite Score",
      Measure == "ADAMS.Total" ~ "ADAMS Total",
      Measure == "PedsQL" ~ "Pediatric Quality of Life"
    )
  )

```

```{r}
#| label: fig-lca-plot

ggplot(test, aes(x = Visit, y = Score, 
                 ymin = seg.min, ymax = seg.max,
                 color = Cluster, group = Cluster)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = 0.1) +
  scale_x_continuous(breaks = c(1, 2, 3)) +
  facet_wrap(.~Measure, nrow = 1, scales = "free_y") +
  geom_blank(aes(y = y_min)) +
  geom_blank(aes(y = y_max)) +
  theme_bw()
```

## DIABLO  

### Results for Tuned Model

```{r, include = FALSE}
# fit diablo without variable selection
k2.diablo <- block.plsda(
  X = X_diff, Y = Y, ncomp = 5, design = design.mat
)

set.seed(9846)
k2.diablo.perf <- perf(
  k2.diablo,
  validation = "loo"
)
# k2.diablo.perf$choice.ncomp

k2.ncomp <- 3
```

```{r}
# Plot Classification by NComp
plot(k2.diablo.perf)
```

Based on the performance plot, the number of components was set to 3.

The number of features to keep per block per component:

<ul>
  <li>Metabolites:
    <ul>
      <li> 20, 10, 10 </li>
    </ul>
  </li>
  <li>Lipids:
    <ul>
      <li> 20, 20, 25 </li>
    </ul>
  </li>
  <li>Peptides:
    <ul>
      <li> 20, 100, 200 </li>
    </ul>
  </li>
  <li>Proteomics:
    <ul>
      <li> 80, 80, 60 </li>
    </ul>
  </li>
</ul>
  

```{r}
plotDIABLO(
  k2.diablo.tune.final,
  ncomp = 1
)
```

```{r}
plotIndiv(
  k2.diablo.tune.final, ind.names = FALSE, legend = TRUE,
  blocks = "average", ellipse = FALSE,
  legend.title = "",
  subtitle = "Multi-Omics Average"
)
```


```{r fig.width=12, fig.height= 8}
cimDiablo(k2.diablo.tune.final, comp = 1, margin = c(8, 20), 
          trim = 3,
          legend.position = "right")
```

```{r include=FALSE}
# Consolidate ROC Curves to a single plot
roc1 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Metabolites",
  print = FALSE
)

roc2 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Lipids",
  print = FALSE
)

roc3 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Peptides",
  print = FALSE
)

roc4 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Proteomics",
  print = FALSE
)

roc <- rbind(
  roc1$graph.Metabolites$comp3$data,
  roc2$graph.Lipids$comp3$data,
  roc3$graph.Peptides$comp3$data,
  roc4$graph.Proteomics$comp3$data
)

roc$Block <- c(
  rep("Metabolites", nrow(roc1$graph.Metabolites$comp3$data)),
  rep("Lipids", nrow(roc2$graph.Lipids$comp3$data)),
  rep("Peptides", nrow(roc3$graph.Peptides$comp3$data)),
  rep("Proteomics", nrow(roc4$graph.Proteomics$comp3$data))
)

roc$Block <- factor(roc$Block, levels = c("Metabolites", "Lipids", "Peptides",
                                          "Proteomics"))
```

```{r}
ggplot(
  roc,
  aes(x = Specificity, y = Sensitivity, color = Block, group = Block)
) +
  geom_line(aes(linetype = Block), linewidth = 1.2, alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0) +
  # add AUC values for each component
  annotate("text", label = "AUC: 0.9185", x = 75, y = 30.1, 
           color = "#F8766D", hjust = 0, vjust = 0) +
  annotate("text", label = "AUC: 0.9481", x = 75, y = 25.1, 
           color = "#7CAE00", hjust = 0, vjust = 0) +
  annotate("text", label = "AUC: 0.9407", x = 75, y = 20.1, 
           color = "#00BFC4", hjust = 0, vjust = 0) +
  annotate("text", label = "AUC: 0.8963", x = 75, y = 15.1, 
           color = "#C77CFF", hjust = 0, vjust = 0) +
  labs(x = "100 - Specificity (%)", y = "Sensitivity (%)") +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  theme_bw()
```

```{r auc-avg}
set.seed(724)
k2.tune.auc <- perf(
  k2.diablo.tune.final,
  auc = TRUE,
  validation = "loo"
)
```

The AUC averaged over all blocks is `r k2.tune.auc$auc$comp3[1]`.

```{r}
## check performance ##
# tuned model
k2.tune.predict <- predict(
  k2.diablo.tune.final,
  newdata = X_diff
)

k2.diablo.tune.confusion <- get.confusion_matrix(
  truth = Y,
  predicted =  k2.tune.predict$WeightedVote$centroids.dist[, k2.ncomp]
)

rownames(k2.diablo.tune.confusion) <- c("Cluster 1", "Cluster 2")

kable(k2.diablo.tune.confusion,
      col.names = c("Predicted Cluster 1", "Predicted Cluster 2"),
      row.names = TRUE) |>
  kable_classic(full_width = FALSE)
```

The tuned DIABLO model has a balanced error rate of `r paste0(round(get.BER(k2.diablo.tune.confusion)*100, 2), "%")`.

# Compare LCA Cluster Assignment to Treatment Group Assignment

Now that the clinical trial has completed, how do our LCA cluster assignments 
compare to our Treatment groups?  

```{r}
compare_clusters <- metformin_lca_2$Data |>
  dplyr::left_join(
    x = _,
    y = unblinding_key |> dplyr::select(fxs_sts_id, unmasked),
    by = c("id" = "fxs_sts_id")
  ) |> 
  dplyr::select(id, cluster, unmasked) |> 
  dplyr::distinct()

compare_clusters |> 
  gtsummary::tbl_cross(
    row = unmasked,
    col = cluster,
    label = list(
      unmasked ~ "Treatment Group",
      cluster ~ "LCA Cluster"
    )
  )
```

## Summarize patient characteristics by 'Subgroup'  

```{r}
subgroup_data <- compare_clusters |> 
  dplyr::left_join(
    x = _,
    y = lca.diablo::redcap_data |> 
      dplyr::filter(redcap_event_name == "baseline_visit_vis_arm_1"),
    by = c("id" = "fxs_sts_id")
  ) |> 
  dplyr::mutate(
    Subgroup = glue::glue("{unmasked}, {cluster}") |> 
      factor(
        levels = c(
          "Metformin, 1", "Metformin, 2",
          "Placebo, 1", "Placebo, 2"
        )
      ),
    Age = demo_age |> as.numeric(),
    tassone_cat = tassone_cat |> toupper(),
    `Age Category` = dplyr::case_when(
      Age < 12 ~ "6-11 yrs",
      Age >= 12 ~ "12+ yrs"
    ) |> 
      factor(levels = c("6-11 yrs", "12+ yrs"))
  )

subgroup_data |> 
  gtsummary::tbl_summary(
    by = Subgroup,
    type = Age ~ "continuous2",
    statistic = Age ~ c("{mean} ({sd})", 
                        "{median} ({p25}, {p75})", 
                        "{min}, {max}"),
    include = c(Age, `Age Category`, tassone_cat)
  ) |> 
  gtsummary::add_overall()
```

## Similarity Network Fusion (SNF) Consensus Subtypes

```{r}
placebo_ids <- compare_clusters |>
  dplyr::filter(unmasked == "Placebo") |>
  dplyr::pull(id) |>
  as.character()

control_ids <- intersect(rownames(X_diff[[1]]), placebo_ids)

snf_fit <- run_snf_subtypes(
  omics = X_diff,
  n_clusters = 2,
  K = 10,
  t = 20,
  control_ids = if (length(control_ids) >= 2) control_ids else NULL
)

snf_annotations <- annotate_snf_subtypes(
  subtypes = snf_fit$subtypes,
  clinical_data = compare_clusters |>
    dplyr::mutate(id = as.character(id)),
  id_col = "id"
) |>
  dplyr::mutate(
    snf_subtype = factor(snf_subtype)
  )

snf_annotations |>
  gtsummary::tbl_cross(
    row = unmasked,
    col = snf_subtype,
    label = list(
      unmasked ~ "Treatment Group",
      snf_subtype ~ "SNF Subtype"
    )
  )
```

```{r}
snf_order <- order(snf_fit$subtypes)
ordered_ids <- rownames(snf_fit$fused_network)[snf_order]
ordered_annotations <- snf_annotations |>
  dplyr::mutate(id = as.character(id)) |>
  dplyr::slice(match(ordered_ids, id))

subtype_palette <- c("Subtype 1" = "#1b9e77", "Subtype 2" = "#d95f02")
treatment_palette <- c("Metformin" = "#7570b3", "Placebo" = "#e7298a")

gplots::heatmap.2(
  snf_fit$fused_network[snf_order, snf_order],
  trace = "none",
  dendrogram = "none",
  Rowv = FALSE,
  Colv = FALSE,
  col = grDevices::colorRampPalette(c("white", "#2c7fb8"))(100),
  RowSideColors = subtype_palette[as.character(ordered_annotations$snf_subtype)],
  ColSideColors = treatment_palette[ordered_annotations$unmasked],
  key.title = "Similarity",
  main = "SNF Fused Similarity Network"
)

legend(
  "topright",
  legend = c(names(subtype_palette), names(treatment_palette)),
  fill = c(subtype_palette, treatment_palette),
  cex = 0.8,
  bty = "n"
)
```

## Heatmaps for Baseline Omics by 'Subgroup'  

```{r}
#| label: process-omics-list
#| include: false

## Visit 1 ##
# metabolomics
metabs <- omics_list$compounds |> 
  tibble::as_tibble(rownames = NA) |> 
  tibble::rownames_to_column("flora_id") |> 
  dplyr::left_join(
    x = _,
    y = patient_design |> dplyr::select(fxs_sts_id, Visit1),
    by = c("flora_id" = "Visit1")
  ) |> 
  dplyr::filter(!is.na(fxs_sts_id)) |> 
  dplyr::left_join(
    x = _,
    y = subgroup_data |> 
      dplyr::select(id, Subgroup),
    by = c("fxs_sts_id" = "id")
  ) |> 
  dplyr::filter(!is.na(Subgroup)) |> 
  tibble::column_to_rownames("fxs_sts_id") |> 
  dplyr::select(-flora_id)

metabs_matrix <- metabs |> dplyr::select(-Subgroup) |> as.matrix()
# lipids
lipids <- omics_list$lipids |> 
  tibble::as_tibble(rownames = NA) |> 
  tibble::rownames_to_column("flora_id") |> 
  dplyr::left_join(
    x = _,
    y = patient_design |> dplyr::select(fxs_sts_id, Visit1),
    by = c("flora_id" = "Visit1")
  ) |> 
  dplyr::filter(!is.na(fxs_sts_id)) |> 
  dplyr::left_join(
    x = _,
    y = subgroup_data |> 
      dplyr::select(id, Subgroup),
    by = c("fxs_sts_id" = "id")
  ) |> 
  dplyr::filter(!is.na(Subgroup)) |> 
  tibble::column_to_rownames("fxs_sts_id") |> 
  dplyr::select(-flora_id)

lipids_matrix <- metabs |> dplyr::select(-Subgroup) |> as.matrix()
# proteins
proteins <- omics_list$proteins |> 
  tibble::as_tibble(rownames = NA) |> 
  tibble::rownames_to_column("flora_id") |> 
  dplyr::left_join(
    x = _,
    y = patient_design |> dplyr::select(fxs_sts_id, Visit1),
    by = c("flora_id" = "Visit1")
  ) |> 
  dplyr::filter(!is.na(fxs_sts_id)) |> 
  dplyr::left_join(
    x = _,
    y = subgroup_data |> 
      dplyr::select(id, Subgroup),
    by = c("fxs_sts_id" = "id")
  ) |> 
  dplyr::filter(!is.na(Subgroup)) |> 
  tibble::column_to_rownames("fxs_sts_id") |> 
  dplyr::select(-flora_id)

proteins_matrix <- metabs |> dplyr::select(-Subgroup) |> as.matrix()
# proteomics
proteomics <- omics_list$proteomics |> 
  tibble::as_tibble(rownames = NA) |> 
  tibble::rownames_to_column("flora_id") |> 
  dplyr::left_join(
    x = _,
    y = patient_design |> dplyr::select(fxs_sts_id, Visit1),
    by = c("flora_id" = "Visit1")
  ) |> 
  dplyr::filter(!is.na(fxs_sts_id)) |> 
  dplyr::left_join(
    x = _,
    y = subgroup_data |> 
      dplyr::select(id, Subgroup),
    by = c("fxs_sts_id" = "id")
  ) |> 
  dplyr::filter(!is.na(Subgroup)) |> 
  tibble::column_to_rownames("fxs_sts_id") |> 
  dplyr::select(-flora_id)

proteomics_matrix <- metabs |> dplyr::select(-Subgroup) |> as.matrix()
```


### Lipids  

```{r}
group_cols <- scales::hue_pal()(4)
row_cols <- lipids |> 
  dplyr::mutate(
    tmp = dplyr::case_when(
      Subgroup == "Metformin, 1" ~ group_cols[1],
      Subgroup == "Metformin, 2" ~ group_cols[2],
      Subgroup == "Placebo, 1" ~ group_cols[3],
      Subgroup == "Placebo, 2" ~ group_cols[4]
    )
  ) |> 
  dplyr::pull(tmp)

gplots::heatmap.2(
  scale(lipids_matrix),
  trace = "none",
  col = gplots::redgreen(100),
  RowSideColors = row_cols
)
```


### Metabolomics

```{r}
group_cols <- scales::hue_pal()(4)
row_cols <- metabs |> 
  dplyr::mutate(
    tmp = dplyr::case_when(
      Subgroup == "Metformin, 1" ~ group_cols[1],
      Subgroup == "Metformin, 2" ~ group_cols[2],
      Subgroup == "Placebo, 1" ~ group_cols[3],
      Subgroup == "Placebo, 2" ~ group_cols[4]
    )
  ) |> 
  dplyr::pull(tmp)

gplots::heatmap.2(
  scale(metabs_matrix),
  trace = "none",
  col = gplots::redgreen(100),
  RowSideColors = row_cols
)
```

### Peptides

```{r}
group_cols <- scales::hue_pal()(4)
row_cols <- proteins |> 
  dplyr::mutate(
    tmp = dplyr::case_when(
      Subgroup == "Metformin, 1" ~ group_cols[1],
      Subgroup == "Metformin, 2" ~ group_cols[2],
      Subgroup == "Placebo, 1" ~ group_cols[3],
      Subgroup == "Placebo, 2" ~ group_cols[4]
    )
  ) |> 
  dplyr::pull(tmp)

gplots::heatmap.2(
  scale(proteins_matrix),
  trace = "none",
  col = gplots::redgreen(100),
  RowSideColors = row_cols
)
```

### Proteomics

```{r}
group_cols <- scales::hue_pal()(4)
row_cols <- proteomics |> 
  dplyr::mutate(
    tmp = dplyr::case_when(
      Subgroup == "Metformin, 1" ~ group_cols[1],
      Subgroup == "Metformin, 2" ~ group_cols[2],
      Subgroup == "Placebo, 1" ~ group_cols[3],
      Subgroup == "Placebo, 2" ~ group_cols[4]
    )
  ) |> 
  dplyr::pull(tmp)

gplots::heatmap.2(
  scale(proteomics_matrix),
  trace = "none",
  col = gplots::redgreen(100),
  RowSideColors = row_cols
)
```
