---
title: "Metformin DIABLO"
date: "`r Sys.Date()`"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    echo: false
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(kableExtra, dplyr, tidyr, lme4, lmerTest, ggplot2, cowplot,
               mixOmics)
devtools::load_all()
```

```{r}
#| label: import-results
#| include: false

## load data ##
X_diff <- readRDS(
  file = "../extdata/archive/DIABLO_X_Difference_list.rds"
)
# update X_diff names
names(X_diff) <- c("Metabolites", "Lipids", "Peptides", "Proteomics")

# remove subj 10
X_diff$Metabolites <- X_diff$Metabolites[-10, ]
X_diff$Lipids <- X_diff$Lipids[-10, ]
X_diff$Peptides <- X_diff$Peptides[-10, ]
X_diff$Proteomics <- X_diff$Proteomics[-10, ]

Y <- readRDS(
  file = paste0(
    "../extdata/archive/k2.Assignments.20230714.rds"
  )
)

design.mat <- readRDS(
  file = "../extdata/archive/Metformin.Difference.Diablo.Design.Matrix.rds"
)

rownames(design.mat) <- colnames(design.mat) <- c(
  "Metabolites", "Lipids", "Peptides", "Proteomics"
)

## load lca results ##
metformin_lca_2 <- readRDS(
  paste0(
    "../extdata/archive/Metformin.k2.LCA.NoOutlier.rds"
  )
)

## load DIABLO results ##
k2.diablo.tune.final <- readRDS(
  paste0(
    "../extdata/archive/k2.DIABLO.Tuned.20230713.rds"
  )
)
```

# Previous Analysis 

## Clustering

```{r avg-response-by-cluster}
lca_2_means <- data.frame(
  "Cluster" = as.factor(c(rep(c(1, 2), each = 3))),
  "Visit" = rep(c(1, 2, 3), 2),
  "ABC.Composite" = c(
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "abc_comp_rev"] * 1,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "abc_comp_rev"] * 2,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "abc_comp_rev"] * 3,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "abc_comp_rev"] * 1,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "abc_comp_rev"] * 2,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "abc_comp_rev"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "abc_comp_rev"] * 3
  ),
  "ADAMS.Total" = c(
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "ADAMS_Total"] * 1,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "ADAMS_Total"] * 2,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "ADAMS_Total"] * 3,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "ADAMS_Total"] * 1,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "ADAMS_Total"] * 2,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "ADAMS_Total"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "ADAMS_Total"] * 3
  ),
  "PedsQL" = c(
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "PedsQL"] * 1,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "PedsQL"] * 2,
    metformin_lca_2$Models$`1`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`1`$Fixedest_mean["visit", "PedsQL"] * 3,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "PedsQL"] * 1,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "PedsQL"] * 2,
    metformin_lca_2$Models$`2`$Fixedest_mean["int", "PedsQL"] + 
      metformin_lca_2$Models$`2`$Fixedest_mean["visit", "PedsQL"] * 3
  )
) |>
  pivot_longer(cols = c(ABC.Composite, ADAMS.Total, PedsQL),
               names_to = "Measure", values_to = "Score") |>
  mutate(
    y_min = 0,
    y_max = case_when(
      Measure == "ABC.Composite" ~ 80,
      Measure == "ADAMS.Total" ~ 40,
      Measure == "PedsQL" ~ 300
    )
  )

# average Std Error
lca_2_ste <- data.frame(
  Cluster = c(1, 2),
  ABC.Composite = c(
    mean(
      c(metformin_lca_2$Models$`1`$Fixedest[[1]]["visit1_", "Std. Error"],
        metformin_lca_2$Models$`1`$Fixedest[[2]]["visit1_", "Std. Error"])
    ),
    mean(
      c(metformin_lca_2$Models$`2`$Fixedest[[1]]["visit1_", "Std. Error"],
        metformin_lca_2$Models$`2`$Fixedest[[2]]["visit1_", "Std. Error"])
    )
  ),
  ADAMS.Total = c(
    mean(
      c(metformin_lca_2$Models$`1`$Fixedest[[1]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`1`$Fixedest[[3]]["visit1_", "Std. Error"])
    ),
    mean(
      c(metformin_lca_2$Models$`2`$Fixedest[[1]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`2`$Fixedest[[3]]["visit1_", "Std. Error"])
    )
  ),
  PedsQL = c(
    mean(
      c(metformin_lca_2$Models$`1`$Fixedest[[2]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`1`$Fixedest[[3]]["visit2_", "Std. Error"])
    ),
    mean(
      c(metformin_lca_2$Models$`2`$Fixedest[[2]]["visit2_", "Std. Error"],
        metformin_lca_2$Models$`2`$Fixedest[[3]]["visit2_", "Std. Error"])
    )
  )
)

# add +/- 1.96*StdError
test <- lca_2_means |>
  mutate(
    seg.min = case_when(
      Measure == "ABC.Composite" & Cluster == 1 ~ Score - (1.96 * lca_2_ste["1", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 1 ~ Score - (1.96 * lca_2_ste["1", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 1 ~ Score - (1.96 * lca_2_ste["1", "PedsQL"]),
      Measure == "ABC.Composite" & Cluster == 2 ~ Score - (1.96 * lca_2_ste["2", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 2 ~ Score - (1.96 * lca_2_ste["2", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 2 ~ Score - (1.96 * lca_2_ste["2", "PedsQL"])
    ),
    seg.max = case_when(
      Measure == "ABC.Composite" & Cluster == 1 ~ Score + (1.96 * lca_2_ste["1", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 1 ~ Score + (1.96 * lca_2_ste["1", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 1 ~ Score + (1.96 * lca_2_ste["1", "PedsQL"]),
      Measure == "ABC.Composite" & Cluster == 2 ~ Score + (1.96 * lca_2_ste["2", "ABC.Composite"]),
      Measure == "ADAMS.Total"  & Cluster == 2 ~ Score + (1.96 * lca_2_ste["2", "ADAMS.Total"]),
      Measure == "PedsQL"  & Cluster == 2 ~ Score + (1.96 * lca_2_ste["2", "PedsQL"])
    )
  )

# rename variables for figures
test <- test |>
  mutate(
    Measure = case_when(
      Measure == "ABC.Composite" ~ "ABC Composite Score",
      Measure == "ADAMS.Total" ~ "ADAMS Total",
      Measure == "PedsQL" ~ "Pediatric Quality of Life"
    )
  )

```

```{r}
#| label: fig-lca-plot

ggplot(test, aes(x = Visit, y = Score, 
                 ymin = seg.min, ymax = seg.max,
                 color = Cluster, group = Cluster)) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = 0.1) +
  scale_x_continuous(breaks = c(1, 2, 3)) +
  facet_wrap(.~Measure, nrow = 1, scales = "free_y") +
  geom_blank(aes(y = y_min)) +
  geom_blank(aes(y = y_max)) +
  theme_bw()
```

## DIABLO  

### Results for Tuned Model

```{r, include = FALSE}
# fit diablo without variable selection
k2.diablo <- block.plsda(
  X = X_diff, Y = Y, ncomp = 5, design = design.mat
)

set.seed(9846)
k2.diablo.perf <- perf(
  k2.diablo,
  validation = "loo"
)
# k2.diablo.perf$choice.ncomp

k2.ncomp <- 3
```

```{r}
# Plot Classification by NComp
plot(k2.diablo.perf)
```

Based on the performance plot, the number of components was set to 3.

The number of features to keep per block per component:

<ul>
  <li>Metabolites:
    <ul>
      <li> 20, 10, 10 </li>
    </ul>
  </li>
  <li>Lipids:
    <ul>
      <li> 20, 20, 25 </li>
    </ul>
  </li>
  <li>Peptides:
    <ul>
      <li> 20, 100, 200 </li>
    </ul>
  </li>
  <li>Proteomics:
    <ul>
      <li> 80, 80, 60 </li>
    </ul>
  </li>
</ul>
  

```{r}
plotDIABLO(
  k2.diablo.tune.final,
  ncomp = 1
)
```

```{r}
plotIndiv(
  k2.diablo.tune.final, ind.names = FALSE, legend = TRUE,
  blocks = "average", ellipse = FALSE,
  legend.title = "",
  subtitle = "Multi-Omics Average"
)
```


```{r fig.width=12, fig.height= 8}
cimDiablo(k2.diablo.tune.final, comp = 1, margin = c(8, 20), 
          trim = 3,
          legend.position = "right")
```

```{r include=FALSE}
# Consolidate ROC Curves to a single plot
roc1 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Metabolites",
  print = FALSE
)

roc2 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Lipids",
  print = FALSE
)

roc3 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Peptides",
  print = FALSE
)

roc4 <- auroc(
  k2.diablo.tune.final,
  roc.comp = k2.ncomp,
  roc.block = "Proteomics",
  print = FALSE
)

roc <- rbind(
  roc1$graph.Metabolites$comp3$data,
  roc2$graph.Lipids$comp3$data,
  roc3$graph.Peptides$comp3$data,
  roc4$graph.Proteomics$comp3$data
)

roc$Block <- c(
  rep("Metabolites", nrow(roc1$graph.Metabolites$comp3$data)),
  rep("Lipids", nrow(roc2$graph.Lipids$comp3$data)),
  rep("Peptides", nrow(roc3$graph.Peptides$comp3$data)),
  rep("Proteomics", nrow(roc4$graph.Proteomics$comp3$data))
)

roc$Block <- factor(roc$Block, levels = c("Metabolites", "Lipids", "Peptides",
                                          "Proteomics"))
```

```{r}
ggplot(
  roc,
  aes(x = Specificity, y = Sensitivity, color = Block, group = Block)
) +
  geom_line(aes(linetype = Block), linewidth = 1.2, alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0) +
  # add AUC values for each component
  annotate("text", label = "AUC: 0.9185", x = 75, y = 30.1, 
           color = "#F8766D", hjust = 0, vjust = 0) +
  annotate("text", label = "AUC: 0.9481", x = 75, y = 25.1, 
           color = "#7CAE00", hjust = 0, vjust = 0) +
  annotate("text", label = "AUC: 0.9407", x = 75, y = 20.1, 
           color = "#00BFC4", hjust = 0, vjust = 0) +
  annotate("text", label = "AUC: 0.8963", x = 75, y = 15.1, 
           color = "#C77CFF", hjust = 0, vjust = 0) +
  labs(x = "100 - Specificity (%)", y = "Sensitivity (%)") +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  theme_bw()
```

```{r auc-avg}
set.seed(724)
k2.tune.auc <- perf(
  k2.diablo.tune.final,
  auc = TRUE,
  validation = "loo"
)
```

The AUC averaged over all blocks is `r k2.tune.auc$auc$comp3[1]`.

```{r}
## check performance ##
# tuned model
k2.tune.predict <- predict(
  k2.diablo.tune.final,
  newdata = X_diff
)

k2.diablo.tune.confusion <- get.confusion_matrix(
  truth = Y,
  predicted =  k2.tune.predict$WeightedVote$centroids.dist[, k2.ncomp]
)

rownames(k2.diablo.tune.confusion) <- c("Cluster 1", "Cluster 2")

kable(k2.diablo.tune.confusion,
      col.names = c("Predicted Cluster 1", "Predicted Cluster 2"),
      row.names = TRUE) |>
  kable_classic(full_width = FALSE)
```

The tuned DIABLO model has a balanced error rate of `r paste0(round(get.BER(k2.diablo.tune.confusion)*100, 2), "%")`.

# Compare LCA Cluster Assignment to Treatment Group Assignment

Now that the clinical trial has completed, how do our LCA cluster assignments 
compare to our Treatment groups?  

```{r}
compare_clusters <- metformin_lca_2$Data |>
  dplyr::left_join(
    x = _,
    y = unblinding_key |> dplyr::select(fxs_sts_id, unmasked),
    by = c("id" = "fxs_sts_id")
  ) |> 
  dplyr::select(id, cluster, unmasked) |> 
  dplyr::distinct()

compare_clusters |> 
  gtsummary::tbl_cross(
    row = unmasked,
    col = cluster,
    label = list(
      unmasked ~ "Treatment Group",
      cluster ~ "LCA Cluster"
    )
  )
```


# Updated DIABLO using Treatment Group instead of LCA Cluster Assignment  

How do the DIABLO results change if we use Treatment Group assignment instead of LCA Cluster assignment?

```{r}
#| label: diablo-tx-group
#| include: false
design_tx <- readRDS(
  "../extdata/archive/mDIABLO_Design_with_Tx.rds"
) |> 
  dplyr::arrange(ID) |> 
  dplyr::filter(ID != 10) |> 
  dplyr::mutate(
    Tx = dplyr::case_when(Treatment.Code == "A" ~ "Metformin",
                          Treatment.Code == "B" ~ "Placebo")
  ) |> 
  dplyr::select(ID, Tx) |> 
  dplyr::distinct()

Y.met <- as.factor(design_tx$Tx)

k2.diablo.met <- block.plsda(
  X = X_diff, Y = Y.met, ncomp = 5, design = design.mat
)

set.seed(5178)
k2.diablo.met.perf <- perf(
  k2.diablo.met,
  validation = "loo"
)

k2.met.ncomp <- 2
```

```{r}
# Plot Classification by NComp
plot(k2.diablo.met.perf)
```

```{r}
#| label: diablo.tuning
#| include: false

# set.seed(618)
# test.keepX <- list(
#   Metabolites = seq(10, 55, 5),
#   Lipids = seq(20, 65, 5),
#   Peptides = seq(20, 200, 20),
#   Proteomics = seq(20, 200, 20)
# )

# DO NOT need to rerun, just changed the block names #
# k2.diablo.tune <- tune.block.splsda(
#   X = X_diff,
#   Y = Y.met,
#   ncomp = k2.met.ncomp,
#   test.keepX = test.keepX,
#   design = design.mat,
#   validation = "loo",
#   BPPARAM = BiocParallel::SnowParam(workers = 4)
# )
# 
# k2.met.keepX <- k2.diablo.tune$choice.keepX
k2.met.keepX <- readRDS("../extdata/k2.met.keepX.rds")

k2.diablo.met.tune.final <- block.splsda(
  X = X_diff,
  Y = Y,
  ncomp = k2.met.ncomp,
  keepX = k2.met.keepX,
  design = design.mat
)
```


```{r}
plotDIABLO(
  k2.diablo.met.tune.final,
  ncomp = 1
)
```

```{r}
plotIndiv(
  k2.diablo.met.tune.final, ind.names = FALSE, legend = TRUE,
  blocks = "average", ellipse = FALSE,
  legend.title = "",
  subtitle = "Multi-Omics Average"
)
```

```{r fig.width=12, fig.height= 8}
#| label: fig-met-cim
#| fig-width: 12
#| fig-height: 8

cimDiablo(k2.diablo.met.tune.final, comp = 1, margin = c(8, 20), 
          trim = 3,
          legend.position = "right")
```

```{r include=FALSE}
# Consolidate ROC Curves to a single plot
roc1.met <- auroc(
  k2.diablo.met.tune.final,
  roc.comp = k2.met.ncomp,
  roc.block = "Metabolites",
  print = FALSE
)

roc2.met <- auroc(
  k2.diablo.met.tune.final,
  roc.comp = k2.met.ncomp,
  roc.block = "Lipids",
  print = FALSE
)

roc3.met <- auroc(
  k2.diablo.met.tune.final,
  roc.comp = k2.met.ncomp,
  roc.block = "Peptides",
  print = FALSE
)

roc4.met <- auroc(
  k2.diablo.met.tune.final,
  roc.comp = k2.met.ncomp,
  roc.block = "Proteomics",
  print = FALSE
)

roc.met <- rbind(
  roc1.met$graph.Metabolites$comp2$data,
  roc2.met$graph.Lipids$comp2$data,
  roc3.met$graph.Peptides$comp2$data,
  roc4.met$graph.Proteomics$comp2$data
)

roc.met$Block <- c(
  rep("Metabolites", nrow(roc1.met$graph.Metabolites$comp2$data)),
  rep("Lipids", nrow(roc2.met$graph.Lipids$comp2$data)),
  rep("Peptides", nrow(roc3.met$graph.Peptides$comp2$data)),
  rep("Proteomics", nrow(roc4.met$graph.Proteomics$comp2$data))
)

roc.met$Block <- factor(roc.met$Block, 
                        levels = c("Metabolites", "Lipids", 
                                   "Peptides", "Proteomics")
                        )
```

```{r}
ggplot(
  roc.met,
  aes(x = Specificity, y = Sensitivity, color = Block, group = Block)
) +
  geom_line(aes(linetype = Block), linewidth = 1.2, alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0) +
  # add AUC values for each component
  annotate("text", 
           label = glue::glue("AUC: {roc1.met$Metabolites$comp2[,'AUC']}"), 
           x = 75, y = 30.1, 
           color = "#F8766D", hjust = 0, vjust = 0) +
  annotate("text", 
           label = glue::glue("AUC: {roc1.met$Lipids$comp2[,'AUC']}"), 
           x = 75, y = 25.1, 
           color = "#7CAE00", hjust = 0, vjust = 0) +
  annotate("text", 
           label = glue::glue("AUC: {roc1.met$Peptides$comp2[,'AUC']}"), 
           x = 75, y = 20.1, 
           color = "#00BFC4", hjust = 0, vjust = 0) +
  annotate("text", 
           label = glue::glue("AUC: {roc1.met$Proteomics$comp2[,'AUC']}"), 
           x = 75, y = 15.1, 
           color = "#C77CFF", hjust = 0, vjust = 0) +
  labs(x = "100 - Specificity (%)", y = "Sensitivity (%)") +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  theme_bw()
```

```{r}
k2.met.tune.auc <- perf(
  k2.diablo.met.tune.final,
  auc = TRUE,
  validation = "loo"
)
```

The AUC averaged over all blocks is `r k2.met.tune.auc$auc$comp2[1]`.

```{r}
## check performance ##
# tuned model
k2.met.tune.predict <- predict(
  k2.diablo.met.tune.final,
  newdata = X_diff
)

k2.diablo.met.tune.confusion.df <- data.frame(
  truth = Y.met,
  predicted =  k2.met.tune.predict$WeightedVote$centroids.dist[, k2.met.ncomp]
) |> 
  dplyr::mutate(
    predicted = ifelse(predicted == "Cluster 1", "Metformin", "Placebo")
  )

k2.diablo.met.tune.confusion.df |> 
  gtsummary::tbl_cross(
    row = truth, col = predicted
  )

Y.met2 <- Y.met |> as.character()
Y.met2[Y.met2 == "Metformin"] <- "Cluster 1"
Y.met2[Y.met2 == "Placebo"] <- "Cluster 2"
Y.met2 <- factor(Y.met2, levels = c("Cluster 1", "Cluster 2"))

k2.diablo.met.tune.confusion <- get.confusion_matrix(
  truth = Y.met2,
  predicted =  k2.met.tune.predict$WeightedVote$centroids.dist[, k2.met.ncomp]
)

rownames(k2.diablo.met.tune.confusion) <- c("Cluster 1", "Cluster 2")

```

The tuned DIABLO model has a balanced error rate of `r paste0(round(get.BER(k2.diablo.met.tune.confusion)*100, 2), "%")`.
